Fast Fourier transformation of electromagnetic data for computationally expensive kernels

### 2021-03-30 ManuscriptTime.tex
Edits

\rtodo 1 \added{\emph{The independence from spatial model complexity is in my
  opinion debatable, because of a potential issue with the accuracy of phase
  information. It’s my experience that a high degree of imaginary-field
  component accuracy is needed for a proper f-t transform. --
  I also needed a high level of numerical accuracy in the f-domain, that is,
  very fine meshes and tight solver tolerances. An aspect which in my opinion
  is not addressed sufficiently in the current manuscript.}: Add somewhere that
  high accurate results are required}.

\rtodo 2 \added{\emph{In my first revision, I actually did not make a request
  for making the literature review "more extensive". Instead, I wrote that the
  authors "should spend more time researching the EM literature that is related
  to the actual topic." As it was and still is, half of the introduction
  (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
  overview. Given the fact that the paper connects directly to Plessix et al.
  (2007) and Mulder et al. (2008), I would have expected a more targeted
  introduction towards the specific advances of this manuscript – instead of
  talking about a choice of only loosely connected and more general topics like
  Lebedev grids, tetrahedral grids, integral equation, direct solvers, etc. I
  like that the introduction is generally concise. My impression remains that
  the introduction can have the same length but be more targeted. Other readers
  may also see it that way. I won’t need another explanation about your
  potentially differing opinion. This is just a suggestion for improvement.
  Feel free to ignore.}}

@@

@@

An important factor in terms of speed and accuracy for time-domain responses
obtained from frequency-domain computations is the selection of the required
frequencies. The fewer frequencies required, the quicker we obtain the
time-domain result. \added{This applies primarily for expensive kernels such
as used in 3D modelling where the computation of the response of a single
frequency takes much longer than the transformation itself.} The same applies
for the reverse operation, the fewer times required, the quicker we obtain
frequency-domain results. Even though we limit our analysis here to
frequency-to-time transformations, most of the arguments are reversible and
applicable to time-to-frequency transformations. We decided to use a regular
spacing of frequencies on a log-scale, rather than an adaptive scheme. This
approach is favourable if a wide range of offsets is needed, as the required
frequencies change with offset and an adaptive frequency selection is therefore
often tailored to a single offset\added{ or a limited range of offset}. Also,
it allows for straightforward parallelization over frequencies, which is not
completely possible with an adaptive scheme.

@@

\rtodo \added{Start this section with the method; fmin, fmax, ndec; below we
interpolate with PCHIP, above we set to zero, in-between we interpolate with
cubic spline.}
\replaced{The actually computed frequencies in our approach are defined by
three key-parameters}{There are three key-parameters that have to be defined}:
\replaced{The m}{M}inimum and maximum frequency ($f_\mr{min}, f_\mr{max}$) and
the number of frequencies per decade ($n_\mr{dec}^f$). \added{This creates a
frequency range which is much smaller than the frequency range required for the
transform.} \replaced{A t}{T}rial-and-error \added{approach} with a 3D code is
very time-intensive. However, a simplified, layered model for the required
survey setup and a fast 1D modeller makes it possible to estimate these
parameters easily. \replaced{An example of this is shown in
Figure~\ref{fig:FFTLog-log}.}{To this end, we created a graphical, interactive
tool using \empymod, as shown in Figure~\ref{fig:FFTLog-log}.}
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{01-FFTLog-log}
  \caption{\replaced{Example}{Interactive} frequency selection for a
    \replaced{simple}{user-provided} layered model (the shown model parameters
    are described in the text). This example shows the impulse response at an
    offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to 10\,Hz with
    five frequencies per decade. \deleted{Note that the error is clipped for
    values smaller than 0.01\,\% and bigger than 100\,\%.}}
  \label{fig:FFTLog-log}
\end{figure}
%
\replaced{It }{The shown example model} is a marine scenario with 1\,km water
depth of resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m
thick target of 100\,\ohmm at 1\,km below the seafloor in a background of
1\,\ohmm. The source is 50\,m above the seafloor, the receiver is on the
seafloor, and the response is the inline $x$-directed $E$-field. The left
subplot shows the imaginary part in the frequency domain and the right plot the
corresponding impulse response in the time domain. The red lines are the
semi-analytical responses. The blue circles indicate the actually computed
responses and the black dots the frequencies which are interpolated or set to
zero. The resulting time-domain response has a relative error of less than
1\,\% everywhere except for very early times.

>>>>

\added{BEGIN TEMPORARY ADDED v v v v}

\rtodo 1 \emph{The independence from spatial model complexity is in my
  opinion debatable, because of a potential issue with the accuracy of phase
  information. It’s my experience that a high degree of imaginary-field
  component accuracy is needed for a proper f-t transform. --
  I also needed a high level of numerical accuracy in the f-domain, that is,
  very fine meshes and tight solver tolerances. An aspect which in my opinion
  is not addressed sufficiently in the current manuscript.}: Add somewhere that
  high accurate results are required.

\rtodo 2 \emph{In my first revision, I actually did not make a request
  for making the literature review "more extensive". Instead, I wrote that the
  authors "should spend more time researching the EM literature that is related
  to the actual topic." As it was and still is, half of the introduction
  (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
  overview. Given the fact that the paper \textbf{connects directly to Plessix
  et al. (2007) and Mulder et al. (2008)}, I would have expected a \textbf{more
  targeted introduction towards the specific advances of this manuscript} –
  instead of talking about a choice of only loosely connected and more general
  topics like Lebedev grids, tetrahedral grids, integral equation, direct
  solvers, etc. I like that the introduction is generally concise. My
  impression remains that the introduction can have the same length but be more
  targeted. Other readers may also see it that way. I won’t need another
  explanation about your potentially differing opinion. This is just a
  suggestion for improvement. Feel free to ignore.}

@@

\added{END TEMPORARY ADDED n n n}

@@

An important factor in terms of speed and accuracy for time-domain responses
obtained from frequency-domain computations is the selection of the required
frequencies. The fewer frequencies required, the quicker we obtain the
time-domain result. \added{This applies primarily for expensive kernels such
as used in 3D modelling where the computation of the response of a single
frequency takes much longer than the transformation itself.} The same applies
for the reverse operation, the fewer times required, the quicker we obtain
frequency-domain results. Even though we limit our analysis here to
frequency-to-time transformations, most of the arguments are reversible and
applicable to time-to-frequency transformations. We decided to use a regular
spacing of frequencies on a log-scale, rather than an adaptive scheme. This
approach is favourable if a wide range of offsets is needed, as the required
frequencies change with offset and an adaptive frequency selection is therefore
often tailored to a single offset\added{ or a limited range of offsets}. Also,
it allows for straightforward parallelization over frequencies, which is not
completely possible with an adaptive scheme.

@@

\hrule

\added{Any Fourier transform needs a certain range of frequencies to obtain
time-domain results. Our approach is not to minimize the required for the
transform, but the number of frequencies for which we actually have to compute
responses. We do this in three zones: Above an upper threshold fmax we set the
frequency-domain response to zero. Below a lower threshold fmin we extrapolate
the frequency-domain response. As we know that the imaginary response goes
towards zero for zero frequency, we can turn it into an interpolation, for
which we use PCHIP. And in-between the thresholds we use cubic spline
interpolation in order to only have to actually compute a few distinct
frequencies. Using this approach we found that computing the responses for
15--25 frequencies is usually good enough to obtain time-domain responses for a
large range of offsets; even though the actual transform might need up to
hundreds of frequencies. The required frequencies depend naturally on the
conductivities and on the chosen acquisition type and layout. A trial-and-error
\dots}

\hrule

\replaced{The actually computed frequencies in our approach are defined by
three key-parameters}{There are three key-parameters that have to be defined}:
\replaced{The m}{M}inimum and maximum frequency ($f_\mr{min}, f_\mr{max}$) and
the number of frequencies per decade ($n_\mr{dec}^f$). \added{This creates a
frequency range which is much smaller than the frequency range required for the
transform.} \replaced{A t}{T}rial-and-error \added{approach} with a 3D code is
very time-intensive. However, a simplified, layered model for the required
survey setup and a fast 1D modeller makes it possible to estimate these
parameters easily. \replaced{An example of this is shown in
Figure~\ref{fig:FFTLog-log}.}{To this end, we created a graphical, interactive
tool using \empymod, as shown in Figure~\ref{fig:FFTLog-log}.}
%
\begin{figure}
  \centering
  \includegraphics[width=\cwidth]{01-FFTLog-log}
  \caption{\replaced{Example}{Interactive} frequency selection for a
    \replaced{simple}{user-provided} layered model (the shown model parameters
    are described in the text). This example shows the impulse response at an
    offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to 10\,Hz with
    five frequencies per decade. \deleted{Note that the error is clipped for
    values smaller than 0.01\,\% and bigger than 100\,\%.}}
  \label{fig:FFTLog-log}
\end{figure}
%
\replaced{It }{The shown example model} is a marine scenario with 1\,km water
depth of resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m
thick target of 100\,\ohmm at 1\,km below the seafloor in a background of
1\,\ohmm. The source is 50\,m above the seafloor, the receiver is on the
seafloor, and the response is the inline $x$-directed $E$-field. The left
subplot shows the imaginary part in the frequency domain and the right plot the
corresponding impulse response in the time domain. The red lines are the
semi-analytical responses. The blue circles indicate the actually computed
responses and the black dots the frequencies which are interpolated or set to
zero. The resulting time-domain response has a relative error of less than
1\,\% everywhere except for very early times.
